Towards Verifying the Bitcoin-S Library

### 2019-06-20 paper/paper.tex
add some before/after code

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] Code.scala:60:3: Stainless doesn't support abstract type members
              type A
              ^^^^^^
\end{lstlisting}

@@

Simplified code.
\begin{lstlisting}[style=scala]
  sealed abstract class CurrencyUnit {
    type A

    def +(c: CurrencyUnit): CurrencyUnit =
      Satoshis(satoshis.underlying + c.satoshis.underlying)

@@

  sealed abstract class Satoshis extends CurrencyUnit {
    override type A = Int64
  }
\end{lstlisting}

@@

@@

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] Code.scala:77:33: Only literal arguments are allowed for BigInt.
              def toBigInt: BigInt = BigInt(toLong)
                                            ^^^^^^
\end{lstlisting}

@@

Again, a simplified code version.
\begin{lstlisting}[style=scala]
  sealed abstract class Satoshis extends CurrencyUnit { 
    def toBigInt: BigInt = BigInt(toLong)
    def toLong: Long = underlying.toLong
  }
\end{lstlisting}

@@

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] Code.scala:7:30: Unknown type parameter type T
             sealed abstract class Number[T <: Number[T]]
                                         ^^^^^^^^^^^^^^
\end{lstlisting}

This is a missing feature in Stainless.
It does not support upper type boundaries on the class itself.
To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.
\begin{lstlisting}[style=scala]
  sealed abstract class Number[T <: Number[T]] extends BasicArithmetic[T]  
\end{lstlisting}

@@

>>>>

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] currency/CurrencyUnits.scala:6:3: Stainless doesn't support abstract type members
             type A
             ^^^^^^
\end{lstlisting}

@@

Before:
\begin{lstlisting}[style=scala]
sealed abstract class CurrencyUnit {
  type A

  def satoshis: Satoshis

  def !=(c: CurrencyUnit): Boolean = !(this == c)

  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis

  def +(c: CurrencyUnit): CurrencyUnit = {
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  }

@@

sealed abstract class Satoshis extends CurrencyUnit {
  override type A = Int64

  override def satoshis: Satoshis = this

  def toBigInt: BigInt = BigInt(toLong)

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}
\end{lstlisting}

@@

After:
\begin{lstlisting}[style=scala]
sealed abstract class CurrencyUnit {
  def satoshis: Satoshis

  def !=(c: CurrencyUnit): Boolean = !(this == c)

  def ==(c: CurrencyUnit): Boolean = satoshis == c.satoshis

  def +(c: CurrencyUnit): CurrencyUnit = {
    Satoshis(satoshis.underlying + c.satoshis.underlying)
  }

  protected def underlying: Int64
}

sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = BigInt(toLong)

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}
\end{lstlisting}

@@

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] currency/CurrencyUnits.scala:26:33: Only literal arguments are allowed for BigInt.
              def toBigInt: BigInt = BigInt(toLong)
                                            ^^^^^^
\end{lstlisting}

@@

Before:
\begin{lstlisting}[style=scala]
sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = BigInt(toLong)

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}
\end{lstlisting}

After:
\begin{lstlisting}[style=scala]
sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = underlying.toBigInt

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}
\end{lstlisting}

@@

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] number/NumberType.scala:8:30: Unknown type parameter type T
  sealed abstract class Number[T <: Number[T]] {
                               ^^^^^^^^^^^^^^

\end{lstlisting}

This is a missing feature in Stainless.
It does not support upper type boundaries on the class itself.
To track this, \href{https://github.com/epfl-lara/stainless/issues/519}{issue 519} was created on GitHub during this work.

Before:
\begin{lstlisting}[style=scala]
sealed abstract class Number[T <: Number[T]] {
  type A = BigInt

  /** The underlying scala number used to to hold the number */
  protected def underlying: A

  def toLong: Long = toBigInt.bigInteger.longValueExact()
  def toBigInt: BigInt = underlying

  /**
    * This is used to determine the valid amount of bytes in a number
    * for instance a UInt8 has an andMask of 0xff
    * a UInt32 has an andMask of 0xffffffff
    */
  def andMask: BigInt

  /** Factory function to create the underlying T, for instance a UInt32 */
  def apply: A => T

  def +(num: T): T = apply(checkResult(underlying + num.underlying))

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): A = {
    require((result & andMask) == result,
      "Result was out of bounds, got: " + result)
    result
  }
}

/**
  * Represents a signed number in our number system
  * Instances of this is [[Int64]]
  */
sealed abstract class SignedNumber[T <: Number[T]] extends Number[T]

/**
  * Represents a int64_t in C
  */
sealed abstract class Int64 extends SignedNumber[Int64] {
  override def apply: A => Int64 = Int64(_)
  override def andMask = 0xffffffffffffffffL
}
\end{lstlisting}

After:
\begin{lstlisting}[style=scala]
sealed abstract class Number {
  type A = BigInt

  /** The underlying scala number used to to hold the number */
  protected def underlying: A

  def toLong: Long = toBigInt.bigInteger.longValueExact()
  def toBigInt: BigInt = underlying

  /**
    * This is used to determine the valid amount of bytes in a number
    * for instance a UInt8 has an andMask of 0xff
    * a UInt32 has an andMask of 0xffffffff
    */
  def andMask: BigInt

  /** Factory function to create the underlying T, for instance a UInt32 */
  def apply: A => Int64

  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): A = {
    require((result & andMask) == result,
      "Result was out of bounds, got: " + result)
    result
  }
}

/**
  * Represents a signed number in our number system
  * Instances of this is [[Int64]]
  */
sealed abstract class SignedNumber extends Number

/**
  * Represents a int64_t in C
  */
sealed abstract class Int64 extends SignedNumber {
  override def apply: A => Int64 = Int64(_)
  override def andMask = 0xffffffffffffffffL
}
\end{lstlisting}

@@

\subsection{Removing toLong}

Stainless output:
\begin{lstlisting}[style=stainless]
  [ Error  ] number/NumberType.scala:14:22: Unknown call to bigInteger on Number.this.toBigInt (BigInt) with arguments List() of type List()
             def toLong: Long = toBigInt.bigInteger.longValueExact()
                                ^^^^^^^^^^^^^^^^^^^
\end{lstlisting}

Now that we do not depend on toLong anymore lets just remove it from out code.

Before:
\begin{lstlisting}[style=scala]
sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = underlying.toBigInt

  def toLong: Long = underlying.toLong

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}  
\end{lstlisting}

After:
\begin{lstlisting}[style=scala]
sealed abstract class Satoshis extends CurrencyUnit {
  override def satoshis: Satoshis = this

  def toBigInt: BigInt = underlying.toBigInt

  def ==(satoshis: Satoshis): Boolean = underlying == satoshis.underlying
}  
\end{lstlisting}

And before:
\begin{lstlisting}[style=scala]
sealed abstract class Number {
  type A = BigInt

  /** The underlying scala number used to to hold the number */
  protected def underlying: A

  def toLong: Long = toBigInt.bigInteger.longValueExact()
  def toBigInt: BigInt = underlying

  /**
    * This is used to determine the valid amount of bytes in a number
    * for instance a UInt8 has an andMask of 0xff
    * a UInt32 has an andMask of 0xffffffff
    */
  def andMask: BigInt

  /** Factory function to create the underlying T, for instance a UInt32 */
  def apply: A => Int64

  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): A = {
    require((result & andMask) == result,
      "Result was out of bounds, got: " + result)
    result
  }
}  
\end{lstlisting}

After:
\begin{lstlisting}[style=scala]
sealed abstract class Number {
  type A = BigInt

  /** The underlying scala number used to to hold the number */
  protected def underlying: A

  def toBigInt: BigInt = underlying

  /**
    * This is used to determine the valid amount of bytes in a number
    * for instance a UInt8 has an andMask of 0xff
    * a UInt32 has an andMask of 0xffffffff
    */
  def andMask: BigInt

  /** Factory function to create the underlying T, for instance a UInt32 */
  def apply: A => Int64

  def +(num: Int64): Int64 = apply(checkResult(underlying + num.underlying))

  /**
    * Checks if the given result is within the range
    * of this number type
    */
  private def checkResult(result: BigInt): A = {
    require((result & andMask) == result,
      "Result was out of bounds, got: " + result)
    result
  }
}
\end{lstlisting}
